---
title: "K2 Project: T1-T2 Behavioral Data"
author: "Angie Wang, Cha Cha Pillai, Mira Guleri"
date: "8/27/2020"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(pastecs)    # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arranging plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling

```

# Background

The K-2 project is a multi-timepoint study with Kindergarteners, 1st graders, and 2nd graders (ages 5 to 8) at Synapse. This report outlines the behavioral data that we collected in Time 1 (in-person) and Time 2 (remote).

N = 43 participants participated in both time points.

| T1 | T2 |
|-|-|
| Oct - Dec 2019 (beginning of school year) | May - Jun 2020 (end of school year) |
| In-person | Remote, via Zoom |
| N = 57 | N = 53 |
| CTOPP Color<br>CTOPP Letter<br>TOWRE (Form A)<br>W-J Letter-Word ID (Form C) | CTOPP Color<br>CTOPP Letter<br>TOWRE (Form B)<br>W-J Letter-Word ID (Form A)<br>CTOPP Elision |
| K2_T1_BehavioralData_20200820.csv | K2_T2_BehavioralData_20200728.csv |

```{r load and clean data, include=FALSE}

# read in data ------------------------------------------
b1_data <- read_csv("K2_T1_BehavioralData_20200820.csv", 
                    col_names = TRUE, 
                    na = "N/A")
b2_data <- read_csv("K2_T2_BehavioralData_20200728.csv", 
                    col_names = TRUE, 
                    na = "N/A")

# clean t1 data -----------------------------------------
b1_clean <- b1_data %>% 
  rename(pid = participant_id, 
         t1.age_year = k2_age,
         t1.age_months = k2_age_months,
         t1.age_days = k2_age_days,
         grade = k2_grade,
         t1.ctopp_c_raw = ctopp_color_raw,
         t1.ctopp_l_raw = ctopp_lttr_raw,
         t1.swe_raw = swe_raw2,
         t1.pde_raw = pde_raw_final,
         t1.swe_age = twre_scaled_age,
         t1.pde_age = pde_scaled_age,
         t1.swe_grade = swe_scaled_grade,
         t1.pde_grade = pde_scaled_grade,
         t1.wj_raw = wj_corr_final) %>% 
  select (pid, t1.age_year, t1.age_months, t1.age_days, grade,
          t1.ctopp_c_raw, t1.ctopp_l_raw,
          t1.swe_raw, t1.pde_raw, t1.swe_age, t1.pde_age,
          t1.swe_grade, t1.pde_grade, t1.wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(t1.age = (t1.age_year + (t1.age_months/12) + (t1.age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  select(-t1.age_months, -t1.age_days) %>% 
  select(pid, grade, t1.age, t1.age_year, everything())

# clean t2 data -----------------------------------------
b2_clean <- b2_data %>% 
  rename(pid = participant_id, 
         t2.age_year = k2_t2_b_age_years,
         t2.age_months = k2_t2_b_age_months,
         t2.age_days = k2_t2_b_age_days,
         grade = k2_t2_b_grade_year,
         t2.ctopp_c_raw = k2_t2_b_color_raw,
         t2.ctopp_l_raw = k2_t2_b_lttr_raw,
         t2.elision_raw = k2_t2_b_elision_raw,
         t2.swe_raw = k2_t2_b_swe_raw_fin,
         t2.pde_raw = k2_t2_b_pde_raw_fin,
         t2.swe_age = k2_t2_b_swe_agescal,
         t2.pde_age = k2_t2_b_pde_agescal,
         t2.swe_grade = k2_t2_b_swe_grdscal,
         t2.pde_grade = k2_t2_b_pde_grdscal,
         t2.wj_raw = k2_t2_b_wj_corr_nb) %>% 
  select (pid, t2.age_year, t2.age_months, t2.age_days, grade,
          t2.ctopp_c_raw, t2.ctopp_l_raw, t2.elision_raw,
          t2.swe_raw, t2.pde_raw, t2.swe_age, t2.pde_age,
          t2.swe_grade, t2.pde_grade, t2.wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(t2.age = (t2.age_year + (t2.age_months/12) + (t2.age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  select(-t2.age_months, -t2.age_days) %>% 
  select(pid, grade, t2.age, t2.age_year, everything())


# join cleaned data from both time points by pid and grade
b_both <- inner_join(b1_clean, b2_clean, by = c("pid", "grade"))

```

## Grade breakdown
```{r grade breakdown, echo=FALSE, message=FALSE}

# make a table of grades
b1_grades <- b1_clean %>% group_by(grade) %>% summarise(T1 = n())
b2_grades <- b2_clean %>% group_by(grade) %>% summarise(T2 = n())
b_all_grades <- b_both %>% group_by(grade) %>% summarise(Both = n()) %>% 
  full_join(b1_grades, by="grade") %>% 
  full_join(b2_grades, by="grade") %>% 
  rename(Grade = grade) %>% 
  select(Grade, T1, T2, Both)

# adding totals to the grade table
b_all_grades <- rbind(b_all_grades, 
                      data.frame(Grade = 'Total', 
                                 T1 = sum(b_all_grades$T1),
                                 T2 = sum(b_all_grades$T2),
                                 Both = sum(b_all_grades$Both)))

# format table for rendered html
kable(b_all_grades) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                position = "left") %>% 
  row_spec(4, bold = T)

```

# Data Cleaning

For each time point: 

* Selected raw scores for CTOPP, TOWRE, W-J, plus scaled scores for TOWRE.
* Did not remove any subjects.

Then combined time points using an inner join by pid.

```{r view combined data, echo=FALSE}

b_both
```

# Descriptive Statistics 

## T1
```{r t1 desc stats, echo=FALSE}

# using stat.desc() from pastecs package

b1_clean_desc <- b1_clean %>% 
  select(- c(pid, grade, t1.age_year)) %>%
  stat.desc(norm = T, p = 0.95) 

# format table for rendered html
kable(b1_clean_desc, 
      digits = 3, # round to 3 digits
      caption = "T1 Descriptive Stats") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"),
                fixed_thead = T,
                font_size = 11)

```

## T2
```{r t2 desc stats, echo=FALSE}

b2_clean_desc <- b2_clean %>% 
  select(- c(pid, grade, t2.age_year)) %>%
  stat.desc(norm = T, p = 0.95)

# format table for rendered html
kable(b2_clean_desc, 
      digits = 3,
      caption = "T2 Descriptive Stats") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                fixed_thead = T,
                font_size = 11)

```

```{r calculate means, echo=FALSE}

# Calculate means by grade, use for plots later

b1_clean_means <- b1_clean %>%
  select(-pid, -t1.age_year) %>% 
  group_by(grade) %>% 
  summarise_all("mean", na.rm=TRUE)

b1_clean_means_K <- b1_clean_means %>% filter(grade == "K")
b1_clean_means_1 <- b1_clean_means %>% filter(grade == "1")
b1_clean_means_2 <- b1_clean_means %>% filter(grade == "2")

b2_clean_means <- b2_clean %>%
  select(-pid, -t2.age_year) %>% 
  group_by(grade) %>% 
  summarise_all("mean", na.rm=TRUE)

b2_clean_means_K <- b2_clean_means %>% filter(grade == "K")
b2_clean_means_1 <- b2_clean_means %>% filter(grade == "1")
b2_clean_means_2 <- b2_clean_means %>% filter(grade == "2")

# TO-DO: Put means by grade in table for rendered file

```

# Visualizations (TODO)

## CTOPP Rapid Color Naming

The Color Naming raw score is the **time (in seconds)** it takes to name all the colors on the page.

```{r ctopp color plots, echo=FALSE}

```

## CTOPP Rapid Letter Naming

The Letter Naming raw score is the **time (in seconds)** it takes to name all the letters on the page.

## CTOPP Elision (T2 Only)

The Elision raw score is the **number of items answered correctly** before reaching the ceiling rule (3 in a row incorrect).

## TOWRE Sight Word Efficiency

The Sight Word Efficiency (SWE) raw score is the **number of words read correctly** in 45 seconds.

## TOWRE Phonemic Decoding Efficiency

The Phonemic Decoding Efficiency (PDE) raw score is the **number of non-words read correctly** in 45 seconds.

## W-J Letter-Word Identification

The W-J raw score is the **number of items answered correctly** before reaching the ceiling rule (6 highest items that were administered are incorrect).
