---
title: 'K-2: T1-T2 visualizations for Synapse audience'
author: "Angie Wang"
date: "11/02/2020"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arranging plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling

# variable naming:
# cc = ctopp color
# cl = ctopp letter
# ce = ctopp elision
# swe = towre swe
# pde = towre pde
# wj = w-j letter-word id

# datafile: t1-t2.csv

```

This report contains the various figures that were used to present data to Synapse (audience: parents and teachers).

<br><br><br>

# Project Background

The K-2 project is a multi-time point study with Kindergarteners, 1st graders, and 2nd graders (ages 5 to 8) at Synapse. This report outlines the behavioral data that we collected in Time 1 (in-person) and Time 2 (remote).

**N = 43** students participated in both time points. Individual assessments may show fewer students due to incomplete or invalid assessments in one or both time points.

| T1 | T2 |
|-|-|
| Oct - Dec 2019 (beginning of school year) | May - Jun 2020 (end of school year) |
| In-person | Remote, via Zoom |

```{r load data, include=FALSE}

# this file contains t1 and t2 data (N=43), with added columns manually coded for rank order grouping

b_both <- read_csv("t1-t2.csv", col_names = TRUE, na = "NA") %>% 
  mutate(grade = fct_relevel(grade,"K", "1", "2")) # reorder factors for plotting

```

```{r t1 desc stats, include=FALSE}

# descriptive stats
b_both_desc <- b_both %>% 
  select(- c(pid, grade, t1.age_year, 
             t1.ctopp_c_raw_group,
             t1.ctopp_l_raw_group,
             t1.swe_raw_group,
             t1.pde_raw_group,
             t1.wj_raw_group)) %>%
  describe() 

# format table for rendered html
kable(b_both_desc, digits = 3) %>%  # round to 3 digits
  kable_styling(bootstrap_options = c("striped", "hover"),
                fixed_thead = T,
                font_size = 11)

```

```{r calculate ctopp items per minute, include = FALSE}

# using this instead of raw score to match with the other tests where "bigger is better"

# estimated items per minute for ctopp scores
# each subtest has 36 items
# 36 items / raw score (sec) * 60 sec = x items

b_both <- b_both %>% 
  mutate(t1.ctopp_c_rate = (36 / t1.ctopp_c_raw * 60),
         t2.ctopp_c_rate = (36 / t2.ctopp_c_raw * 60),
         t1.ctopp_l_rate = (36 / t1.ctopp_l_raw * 60),
         t2.ctopp_l_rate = (36 / t2.ctopp_l_raw * 60))

```

```{r calculate towre items per minute, include = FALSE}

# estimated items per minute for towre scores
# raw score / 45 sec * 60 sec = x items

b_both <- b_both %>% 
  mutate(t1.swe_rate = (t1.swe_raw / 45 * 60),
         t2.swe_rate = (t2.swe_raw / 45 * 60),
         t1.pde_rate = (t1.pde_raw / 45 * 60),
         t2.pde_rate = (t2.pde_raw / 45 * 60))

```

```{r calculate time point differences, include=FALSE}

# calculate score differences between time points

b_both <- b_both %>% 
  mutate(ctopp_c_diff = (t2.ctopp_c_rate - t1.ctopp_c_rate),
         ctopp_l_diff = (t2.ctopp_l_rate - t1.ctopp_l_rate),
         swe_diff = (t2.swe_rate - t1.swe_rate),
         pde_diff = (t2.pde_rate - t1.pde_rate),
         wj_diff = (t2.wj_raw - t1.wj_raw))

```


<br><br><br>


# Measuring Growth Over the School Year

*How are things changing over the school year at Synapse?*

<br><br>

The following figures show the relationship between scores in T1 and T2. 

For each of the scatterplots: 

* **Each point represents a group of students**, color-coded by grade. Students were first grouped according to how their T1 score ranked within their grade, yielding three groups per grade. T1 and T2 raw scores were then averaged within each group.

* The T1 score is on the x-axis.

* Growth, the difference between T1 and T2 scores (T2 minus T1), is on the y-axis.* 

* The solid black line at y = 0 represents no change between T1 and T2 scores. Points lying above this line represent improvement in T2 scores. The distance between a data point and this line represents the magnitude of change.
  
* The grey line is the linear regression line, which is the best-fitting line through the data points**, and represents a prediction of T2 change scores based on scores from T1. The vertical distance between a point and the regression line represents the errors of prediction.

<br>

Footnotes:

*This is the difference between the T1 and T2 group averages.

**The regression line is fitted on non-aggregated data (i.e., prior to grouping by T1 score rank). 

<br><br><br>

## W-J Letter-Word Identification

The Letter-Word Identification raw score is the **number of items answered correctly** before reaching the ceiling rule (6 highest items that were administered are incorrect).

This figure shows the performance of students who completed the assessment at both time points (N = 43). Students were grouped based on their T1 raw score rank within their grade.

```{r w-j scatterplot grouped, echo=FALSE, message=F, warning=F, fig.width=8.5, fig.height=8}

# group data and average based on t1 rank
wj_grouped <- b_both %>% 
  select(grade, t1.wj_raw, t1.wj_raw_group, t2.wj_raw) %>% 
  group_by(grade, t1.wj_raw_group) %>% 
  summarise(t1.wj_groupavg = mean(t1.wj_raw),
            t2.wj_groupavg = mean(t2.wj_raw)) %>% 
  mutate(diff.wj_groupavg = (t2.wj_groupavg - t1.wj_groupavg))


# plot scatter based on grouped data
p_wj_scatter <- wj_grouped %>% 
  drop_na(t1.wj_groupavg, diff.wj_groupavg) %>% 
  ggplot(aes(x = t1.wj_groupavg, y = diff.wj_groupavg, color = grade)) + 
  
  # y = 0 and x = 0 lines
  geom_abline(slope = 0, intercept = 0) + 
  geom_vline(xintercept = 0) +
  
  # regression line on ungrouped data
  geom_smooth(aes(x = t1.wj_raw, y = wj_diff), 
              data = b_both, 
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) + 
  
  # scatterplot
  geom_point(size = 10) +
  
  # adjust appearance
  xlim(0, 70) +
  labs(title = "Letter-Word Identification (N = 43, grouped)", 
         x = "T1 Score (# of correct items)",
         y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_wj_scatter
#ggsave("wj_scatter.png")

```

Number of students per group:

| Grade | 1st Tercile | 2nd Tercile | 3rd Tercile |
|-|-|-|-|
| K | 4 | 4 | 5 |
| 1 | 4 | 4 | 5 |
| 2 | 6 | 6 | 5 |

<br><br><br>

## TOWRE

### Sight Word Efficiency

The Sight Word Efficiency (SWE) score below is the **number of words read correctly** in 45 seconds.

This figure shows the performance of students who completed the assessment at both time points (N = 42). Students were grouped based on their T1 raw score rank within their grade.

```{r swe scatterplot grouped, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# group data and average based on t1 rank--------------------------------------
swe_grouped <- b_both %>% 
  select(grade, t1.swe_rate, t1.swe_raw_group, t2.swe_rate) %>% 
  group_by(grade, t1.swe_raw_group) %>% 
  summarise(t1.swe_groupavg = mean(t1.swe_rate),
            t2.swe_groupavg = mean(t2.swe_rate)) %>% 
  mutate(diff.swe_groupavg = (t2.swe_groupavg - t1.swe_groupavg))


# plot scatter of grouped data-------------------------------------------------
p_swe_scatter <- swe_grouped %>% 
  drop_na(t1.swe_groupavg, diff.swe_groupavg) %>% 
  ggplot(aes(x = t1.swe_groupavg, y = diff.swe_groupavg, color = grade)) +
  
  # y = 0 and x = 0 lines
  geom_abline(slope = 0, intercept = 0) +
  geom_vline(xintercept = 0) +
  
  # plot age norms on t1 x-axis
  #geom_vline(xintercept=38.67, color="grey", linetype="dashed") + # 6 y/o norm wpm
  #geom_vline(xintercept=62.67, color="grey", linetype="dashed") + # 7 y/o norm wpm
  #geom_vline(xintercept=73.33, color="grey", linetype="dashed") + # 8 y/o norm wpm
  
  # regression line on ungrouped data
  geom_smooth(aes(x = t1.swe_rate, y = swe_diff), 
              data = b_both, 
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) + 
  
  # scatterplot
  geom_point(size = 10) +
  
  # adjust appearance
  xlim(0, 100) +
  ylim(0, 30) +
  labs(title = "Sight Word Efficiency (N = 42, grouped)", 
       x = "T1 Score (# words per minute)",
       y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_swe_scatter
#ggsave("swe_scatter.png")

```

Number of students per group:

| Grade | 1st Tercile | 2nd Tercile | 3rd Tercile |
|-|-|-|-|
| K | 4 | 4 | 4 |
| 1 | 4 | 4 | 5 |
| 2 | 6 | 6 | 5 |

<br><br><br>

### Phonemic Decoding Efficiency

The Phonemic Decoding Efficiency (PDE) raw score is the **number of non-words read correctly** in 45 seconds.

This figure shows the performance of students who completed the assessment at both time points (N = 38). Students were grouped based on their T1 raw score rank within their grade.

```{r pde scatterplot grouped, echo=F, warning=F, message=F, fig.width=8.5, fig.height=8}

# group data and average based on t1 rank----------------------------
pde_grouped <- b_both %>% 
  select(grade, t1.pde_rate, t1.pde_raw_group, t2.pde_rate) %>% 
  group_by(grade, t1.pde_raw_group) %>% 
  summarise(t1.pde_groupavg = mean(t1.pde_rate),
            t2.pde_groupavg = mean(t2.pde_rate)) %>% 
  mutate(diff.pde_groupavg = (t2.pde_groupavg - t1.pde_groupavg))

# plot scatter of grouped data---------------------------------------
p_pde_scatter <- pde_grouped %>% 
  drop_na(t1.pde_groupavg, diff.pde_groupavg) %>% 
  ggplot(aes(x = t1.pde_groupavg, y = diff.pde_groupavg, color = grade)) + 
  
  # y = 0 and x = 0 lines
  geom_abline(slope = 0, intercept = 0) +
  geom_vline(xintercept = 0) +
  
  # regression line on ungrouped data
  geom_smooth(aes(x = t1.pde_rate, y = pde_diff), 
              data = b_both, 
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) + 
  
  # scatterplot
  geom_point(size = 10) +
  
  # adjust appearance
  ylim(-6, 30) +
  labs(title = "Phonemic Decoding Efficiency (N = 38, grouped)", 
       x = "T1 Score (# nonwords per minute)",
       y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_pde_scatter
#ggsave("pde_scatter.png")

```

Number of students per group:

| Grade | 1st Tercile | 2nd Tercile | 3rd Tercile |
|-|-|-|-|
| K | 3 | 3 | 3 |
| 1 | 4 | 4 | 5 |
| 2 | 5 | 5 | 6 |


```{r pde scatterplot disaggregated, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# plot scatter of ungrouped data------------------------------------
p_pde_ungrouped <- b_both %>% 
  drop_na(t1.pde_rate, pde_diff) %>% 
  ggplot(aes(x = t1.pde_rate, y = pde_diff, color = grade)) + 
  
  # y = 0 and x = 0 lines
  geom_abline(slope = 0, intercept = 0) +
  geom_vline(xintercept = 0) +
  
  # regression line on ungrouped data
  geom_smooth(aes(x = t1.pde_rate, y = pde_diff, color = grade), 
              data = b_both, 
              method=  lm, 
              se = F,
              fullrange = F) + 
  
  # scatterplot showing individual subs
  geom_point(size = 1) +
  
  # adjust appearance
  labs(title = "Phonemic Decoding Efficiency (N = 38, ungrouped)", 
         x = "T1 Score (# nonwords per minute)",
         y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_pde_ungrouped
#ggsave("pde_scatter_ungrouped.png")

```

<br><br><br>

### Overall

Each point in the dotplot represents the average subtest score by time point (split by grade).

```{r towre overall dotplot, echo=F, message=F, warning=F, fig.width=8, fig.height=5}

# arrange data----------------------------------------------------------------
towre_overall <- b_both %>% 
  select(grade, t1.swe_rate, t2.swe_rate, t1.pde_rate, t2.pde_rate) %>%
  group_by(grade) %>% 
  summarise(T1.SWE = mean(t1.swe_rate, na.rm = TRUE),
            T2.SWE = mean(t2.swe_rate, na.rm = TRUE),
            T1.PDE = mean(t1.pde_rate, na.rm = TRUE),
            T2.PDE = mean(t2.pde_rate, na.rm = TRUE)) %>% 
  gather(key = "key", value = "mean", T1.SWE:T2.PDE) %>% 
  separate(key, c("time", "subtest"))


# plot dotplot------------------------------------------------
p_towre_overall <- ggplot(towre_overall, aes(x = time, y = mean, fill = subtest)) + 
  
  geom_dotplot(binaxis = 'y', stackdir='center', dotsize = 2) +
  facet_grid(. ~ grade) +
  geom_line(aes(group = subtest), 
            arrow = arrow(angle=30, length = unit(0.13, "cm"), type="open")) +
  
  labs(title = "Sight Word Efficiency (SWE) and Phonemic Decoding Efficiency (PDE)",
       x = "Time Point",
       y = "Score (items per minute)") +
  
  theme_light() +
  theme(strip.text.x = element_text(size=15, color="black", face="bold")) +

  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000")) # synapse colors


p_towre_overall
#ggsave("towre_overall.png", p_towre_overall)

```

<br><br><br>

## CTOPP RAN

### Rapid Color Naming

The Color Naming score below is the **estimated number of items named per minute**.

This figure shows the performance of students who completed the assessment at both time points (N = 43). Students were grouped based on their T1 score rank within their grade. 

```{r color scatterplot grouped, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# group data and average based on t1 rank------------------------------------------
cc_grouped <- b_both %>% 
  select(grade, t1.ctopp_c_rate, t1.ctopp_c_raw_group, t2.ctopp_c_rate) %>% 
  group_by(grade, t1.ctopp_c_raw_group) %>% 
  summarise(t1.ctopp_c_groupavg = mean(t1.ctopp_c_rate),
            t2.ctopp_c_groupavg = mean(t2.ctopp_c_rate)) %>% 
  mutate(diff.ctopp_c_groupavg = (t2.ctopp_c_groupavg - t1.ctopp_c_groupavg))


# plot scatter of grouped data----------------------------------------------------
p_cc_scatter <- ggplot(cc_grouped, aes(x = t1.ctopp_c_groupavg, 
                                  y = diff.ctopp_c_groupavg)) + 
  
  # y = 0 line
  geom_abline(slope = 0, intercept = 0) +
  
  # regression line on ungrouped data 
  geom_smooth(aes(x = t1.ctopp_c_rate, y = ctopp_c_diff), 
              data = b_both, 
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) +
  
  # scatterplot grouped
  geom_point(aes(color = grade), size = 10) +
  
  # adjust appearance
  ylim(-15, 30) +
  labs(title = "Rapid Color Naming (N = 43, grouped)", 
       x = "T1 Score (items per minute)",
       y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000")) # synapse colors

p_cc_scatter
#ggsave("color_scatter.png")

```

Number of students per group:

| Grade | 1st Tercile | 2nd Tercile | 3rd Tercile |
|-|-|-|-|
| K | 4 | 4 | 5 |
| 1 | 4 | 4 | 5 |
| 2 | 6 | 6 | 5 |


<br><br><br>

### Rapid Letter Naming

The Letter Naming score below is the **estimated number of items named per minute**.

This figure shows the performance of students who completed the assessment at both time points (N = 42). Students were grouped based on their T1 raw score rank within their grade.

```{r letter scatterplot grouped, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# group data and average based on t1 rank-------------------------------------
cl_grouped <- b_both %>% 
  select(grade, t1.ctopp_l_rate, t1.ctopp_l_raw_group, t2.ctopp_l_rate) %>% 
  group_by(grade, t1.ctopp_l_raw_group) %>% 
  summarise(t1.ctopp_l_groupavg = mean(t1.ctopp_l_rate),
            t2.ctopp_l_groupavg = mean(t2.ctopp_l_rate)) %>% 
  mutate(diff.ctopp_l_groupavg = (t2.ctopp_l_groupavg - t1.ctopp_l_groupavg))


# plot scatter of grouped data------------------------------------------------
p_cl_scatter <- cl_grouped %>% 
  drop_na(t1.ctopp_l_groupavg, diff.ctopp_l_groupavg) %>%
  ggplot(aes(x = t1.ctopp_l_groupavg, y = diff.ctopp_l_groupavg, color = grade)) + 
  
  # y = 0 line
  geom_abline(slope = 0, intercept = 0) +
 
  # regression line on ungrouped data 
  geom_smooth(aes(x = t1.ctopp_l_rate, y = ctopp_l_diff), 
              data = b_both, 
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) + 
  
  # scatterplot grouped
  geom_point(size = 10) +
  
  # adjust appearance
  ylim(-15, 30) +
  labs(title = "Rapid Letter Naming (N = 42, grouped)", 
       x = "T1 Score (items per minute)",
       y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000")) # synapse colors

p_cl_scatter
#ggsave("letter_scatter.png")

```

Number of students per group:

| Grade | 1st Tercile | 2nd Tercile | 3rd Tercile |
|-|-|-|-|
| K | 4 | 4 | 4 |
| 1 | 4 | 4 | 5 |
| 2 | 6 | 6 | 5 |

<br><br><br>

### Overall

```{r ctopp combined scatter, echo=FALSE, message=F, fig.width=8.5, fig.height=8}

# combining and arranging the data-------------------------------------------
ctopp_both <- cl_grouped %>% 
  drop_na() %>% 
  rename(group = t1.ctopp_l_raw_group)

ctopp_both <- cc_grouped %>% 
  rename(group = t1.ctopp_c_raw_group) %>% 
  inner_join(ctopp_both, by = "group") %>% 
  rename(grade = grade.x) %>% 
  select(-grade.y)


# plot scatter of grouped data------------------------------------------------
p_ran_scatter <- ggplot(ctopp_both) +
  
  # y = 0 line
  geom_abline(slope = 0, intercept = 0) +
  
  # letter: scatterplot of grouped data
  geom_point(aes(x = t1.ctopp_l_groupavg, y = diff.ctopp_l_groupavg, color = grade), 
             size = 10, alpha = 0.7) + 
  
  # color: scatterplot of grouped data
  geom_point(aes(x = t1.ctopp_c_groupavg, y = diff.ctopp_c_groupavg, color = grade),
             size = 10, alpha = 0.3, show.legend = F) + 
  
  # add lines between color and letter points
  geom_segment(aes(x = t1.ctopp_c_groupavg , y = diff.ctopp_c_groupavg, 
                   xend = t1.ctopp_l_groupavg , yend = diff.ctopp_l_groupavg, 
                   color=grade), 
               arrow = arrow(angle=30, length = unit(0.1, "inches"), type="closed"), 
               show.legend = F) +
  
  # adjust appearance
  ylim(-15, 30) +
  labs(title = "Rapid Naming: Color and Letter", 
       x = "T1 Score (items per minute)",
       y = "Growth (T2 - T1)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000")) # synapse colors

p_ran_scatter
#ggsave("ctopp_scatter.png")

```

```{r ctopp combined dotplot, echo=FALSE, message=F, fig.width=8, fig.height=5}

# arrange data----------------------------------------------------------------
ctopp_both_nogroup <- b_both %>% 
  select(grade, t1.ctopp_c_rate, t2.ctopp_c_rate, 
         t1.ctopp_l_rate, t2.ctopp_l_rate) %>%
  group_by(grade) %>% 
  summarise(T1.Color = mean(t1.ctopp_c_rate, na.rm = TRUE),
            T2.Color = mean(t2.ctopp_c_rate, na.rm = TRUE),
            T1.Letter = mean(t1.ctopp_l_rate, na.rm = TRUE),
            T2.Letter = mean(t2.ctopp_l_rate, na.rm = TRUE)) %>% 
  gather(key = "key", value = "mean", T1.Color:T2.Letter) %>% 
  separate(key, c("time", "subtest"))


# plot dotplot of grouped data------------------------------------------------
p_ran_dotplot <- ggplot(ctopp_both_nogroup, aes(x = time, y = mean, fill = subtest)) + 
  
  # plot dotplot
  geom_dotplot(binaxis = 'y', stackdir='center', dotsize = 2) +
  
  # split into panels by grade
  facet_grid(. ~ grade) +
  
  # add lines connecting t1-t2
  geom_line(aes(group = subtest), 
            arrow = arrow(angle=30, length = unit(0.13, "cm"), type="open")) +
  
  # adjust appearance
  labs(title = "Rapid Naming: Color and Letter",
       x = "Time Point",
       y = "Score (items per minute)") +
  theme_light() +
  theme(strip.text.x = element_text(size=15, color="black", face="bold")) +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000")) # synapse colors

p_ran_dotplot
#ggsave("ctopp_dotplot.png")

```

<br><br><br>


# Correlation Structure

*How are the various measures related?*

<br><br>


## Overall Correlation between T1 and T2 Measures

There is a strong correlation between T1 and T2 for each measure (r ~= 0.9, except for Rapid Color Naming). 

```{r overall correlations, echo=F}

b_both_corr <- b_both %>% 
  mutate(t1.ctopp_c_flip = -1 * t1.ctopp_c_raw,
         t1.ctopp_l_flip = -1 * t1.ctopp_l_raw,
         t2.ctopp_c_flip = -1 * t2.ctopp_c_raw,
         t2.ctopp_l_flip = -1 * t2.ctopp_l_raw) %>% 
  select(t1.ctopp_c_flip, t1.ctopp_l_flip, t1.swe_raw, t1.pde_raw, t1.wj_raw,
         t2.ctopp_c_flip, t2.ctopp_l_flip, t2.swe_raw, t2.pde_raw, t2.wj_raw) %>% 
  cor(use = "na.or.complete")
  

colnames(b_both_corr) <- c("T1 Color", "T1 Letter", "T1 SWE", "T1 PDE", "T1 LWID",
                           "T2 Color", "T2 Letter", "T2 SWE", "T2 PDE", "T2 LWID")
rownames(b_both_corr) <- c("T1 Color", "T1 Letter", "T1 SWE", "T1 PDE", "T1 LWID",
                           "T2 Color", "T2 Letter", "T2 SWE", "T2 PDE", "T2 LWID")

#pdf(file="overallCorrelations.pdf")
overall_corr_circle <- corrplot(b_both_corr[c(1, 2, 3, 4, 5), 1:10], method = "circle", type="upper", tl.col = "black", tl.srt = 45)
overall_corr_number <- corrplot(b_both_corr[c(1, 2, 3, 4, 5), 1:10], method = "number", type="upper", tl.col = "black", tl.srt = 45)
#dev.off()

```


## Growth Correlations

For each subject, the "growth" value is the difference between T1 and T2 raw scores (difference = T1 - T2 for raw scores that are "time to finish"; difference = T2 - T1 for raw scores that are "number correct"). 

The following figures show correlations of these differences across the five measures that were administered in both time points.

```{r correlations between time points, echo=F}

b_diff <- b_both %>% 
  select(grade, ctopp_c_diff, ctopp_l_diff, swe_diff, pde_diff, wj_diff)

# compute correlation on differences
b_diff_corr <- b_diff %>% 
  select(-grade) %>% 
  cor(use = "na.or.complete")

colnames(b_diff_corr) <- c("Color growth", "Letter growth", "SWE growth",
                           "PDE growth", "LWID growth")
rownames(b_diff_corr) <- c("Color growth", "Letter growth", "SWE growth",
                           "PDE growth", "LWID growth")

# plot correlation matrix
#pdf(file="growthCorrelations.pdf")
growth_corr_circle <- corrplot(b_diff_corr, method="circle", type="upper", tl.col="black", tl.srt=45)
growth_corr_number <- corrplot(b_diff_corr, method="number", type="upper", tl.col="black", tl.srt=45)
#dev.off()
    
```


### By grade

#### Kindergarten

```{r correlations between time points, kindergarten, echo=F}

# same thing as above but only with kindergartners
b_diff_corr_K <- b_diff %>% 
  filter(grade == "K") %>% 
  select(- grade) %>% 
  cor(use = "na.or.complete")

# plot side by side
par(mfrow=c(1,2))
p_corr_gk_1 <- corrplot(b_diff_corr_K, method = "color")
p_corr_gk_2 <- corrplot(b_diff_corr_K, method = "number")
    
```

#### First Grade

```{r correlations between time points, first grade, echo=F}

# same thing as above but only with first graders
b_diff_corr_1 <- b_diff %>% 
  filter(grade == "1") %>% 
  select(- grade) %>% 
  cor(use = "na.or.complete")

# plot side by side
par(mfrow=c(1,2))
p_corr_g1_1 = corrplot(b_diff_corr_1, method = "color")
p_corr_g1_2 = corrplot(b_diff_corr_1, method = "number")
    
```

#### Second Grade

```{r correlations between time points, second grade, echo=F}

# same thing as above but only with second graders
b_diff_corr_2 <- b_diff %>% 
  filter(grade == "2") %>% 
  select(- grade) %>% 
  cor(use = "na.or.complete")

# plot side by side
par(mfrow=c(1,2))
p_corr_g2_1 = corrplot(b_diff_corr_2, method = "color")
p_corr_g2_2 = corrplot(b_diff_corr_2, method = "number")
    
```


## LWID and SWE Growth Correlation

```{r lwid and swe correlation by individuals, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# plot scatter with individual points---------------------------------------
p_wj_swe_ungrouped <- ggplot(b_both, aes(x = wj_diff, y = swe_diff)) + 
  
  # y = 0 and x = 0 lines
  geom_vline(xintercept=0) +
  geom_abline(slope = 0, intercept = 0) +

  # regression line by grade
  geom_smooth(aes(x = wj_diff, y = swe_diff, color = grade), 
              method=  lm, 
              se = F,
              fullrange = F) + 
  
  # scatterplot individual points
  geom_point(aes(color = grade), size = 2) +
  
  # adjust appearance
  xlim(-5, 27) +
  ylim(-5, 35) +
  labs(title = "Relationship between growth on LWID and growth on SWE (N = 42)", 
       x = "LWID Growth (T2 change in # words read)",
       y = "SWE Growth (T2 change in # words read per minute)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_wj_swe_ungrouped
#ggsave("wj_swe_growthCorrelation.png")

```

```{r lwid and swe correlation by group, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

b_both_diff <- read.csv("t1-t2_groupedByDiff.csv") %>% 
  mutate(grade = fct_relevel(grade,"K", "1", "2")) # reorder factors for plotting

# group data and average based on t1 rank--------------------------------------
wj_swe_group_grade <- b_both_diff %>% 
  select(grade, wj_diff, wj_diff_group, swe_diff) %>% 
  drop_na() %>% 
  group_by(grade, wj_diff_group) %>% 
  summarise(wj_diff_groupavg = mean(wj_diff),
            swe_diff_groupavg = mean(swe_diff))

# plot scatter with grouped points-----------------------------------------------
p_wj_swe_grouped <- ggplot(wj_swe_group_grade, aes(x = wj_diff_groupavg, y = swe_diff_groupavg)) + 
 
  # y = 0 and x = 0 lines
  geom_vline(xintercept=0) +
  geom_abline(slope = 0, intercept = 0) +

  # regression line on ungrouped data
  geom_smooth(aes(x = wj_diff, y = swe_diff), 
              data = b_both_diff,
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) + 
  
  # scatterplot grouped
  geom_point(aes(color = grade), size = 10) +
  
  # adjust appearance
  ylim(0, 30) +
  labs(title = "Relationship between growth on LWID and growth on SWE (N = 42, grouped)", 
       x = "LWID Growth (T2 change in # words read)",
       y = "SWE Growth (T2 change in # words per minute)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_wj_swe_grouped
#ggsave("wj_swe_growthCorrelationGrouped.png")

```

## Letter and SWE Correlation

```{r letter and swe correlation by individuals, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# plot scatter with individual points------------------------------------
p_cl_swe_ungrouped <- ggplot(b_both, aes(x = ctopp_l_diff, y = swe_diff)) +
  
  # y = 0 and x = 0 lines
  geom_vline(xintercept=0) +
  geom_abline(slope = 0, intercept = 0) +

  # regression line by grade
  geom_smooth(aes(x = ctopp_l_diff, y = swe_diff, color = grade), 
              method=  lm, 
              se = F,
              fullrange = F) +
  
  # scatterplot of individual points
  geom_point(aes(color = grade), size = 2) +

  # adjust appearance
  labs(title = "Relationship between growth on Letter and growth on SWE (N = 42)", 
       x = "Letter Growth (T2 change in # letters per minute)",
       y = "SWE Growth (T2 change in # words per minute)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_cl_swe_ungrouped
#ggsave("letter_swe_growthCorrelation.png")

```

```{r letter and swe correlation by group, echo=F, message=F, warning=F, fig.width=8.5, fig.height=8}

# group data and average based on t1 rank-----------------------------------
cl_swe_group_grade <- b_both_diff %>% 
  select(grade, ctopp_l_diff, ctopp_l_diff_group, swe_diff) %>% 
  drop_na() %>% 
  group_by(grade, ctopp_l_diff_group) %>% 
  summarise(ctopp_l_diff_groupavg = mean(ctopp_l_diff),
            swe_diff_groupavg = mean(swe_diff))

# plot scatter with grouped points------------------------------------------
p_cl_swe_grouped <- ggplot(cl_swe_group_grade, aes(x = ctopp_l_diff_groupavg, y = swe_diff_groupavg)) + 
  
  # y = 0 and x = 0 lines
  geom_vline(xintercept=0) +
  geom_abline(slope = 0, intercept = 0) +

  # regression line on ungrouped data
  geom_smooth(aes(x = ctopp_l_diff, y = swe_diff), 
              data = b_both_diff,
              method=  lm, 
              se = F,
              color = "grey",
              fullrange = T) +
  
  # scatterplot grouped
  geom_point(aes(color = grade), size = 10) +

  # adjust appearance
  labs(title = "Relationship between growth on Letter and growth on SWE (N = 42, grouped)", 
       x = "Letter Growth (T2 change in # letters per minute)",
       y = "SWE Growth (T2 change in # words read per minute)") +
  theme_light() +
  scale_color_manual(values=c("#007dba", "#9ea2a2", "#000000"))

p_cl_swe_grouped
#ggsave("letter_swe_growthCorrelation_grouped.png")

```