---
title: "K2 Project: T1-T2 Behavioral Data"
author: "Angie Wang, Cha Cha Pillai, Mira Guleri"
date: "9/23/2020"
output: 
  html_document:
    theme: spacelab
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# data files you'll need:
# K2_T1_BehavioralData_20200820.csv
# K2_T2_BehavioralData_20200901.csv

# load libraries
library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(pastecs)    # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(DT)         # for rendering interactive html tables
library(ggpubr)     # for arranging plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling

# knitr: knit to html to avoid the need need to install TeX; you can still 
# save as a PDF in html by using Print webpage -> Save as PDF.

# variable naming:
# cc = ctopp color
# cl = ctopp letter
# ce = ctopp elision
# swe = towre swe
# pde = towre pde
# wj = w-j letter-word id

```

# Background

The K-2 project is a multi-time point study with Kindergarteners, 1st graders, and 2nd graders (ages 5 to 8) at Synapse. This report outlines the behavioral data that we collected in Time 1 (in-person) and Time 2 (remote).

**N = 43** participants participated in both time points.

| T1 | T2 |
|-|-|
| Oct - Dec 2019 (beginning of school year) | May - Jun 2020 (end of school year) |
| In-person | Remote, via Zoom |
| N = 57 | N = 53 |
| CTOPP Color<br>CTOPP Letter<br>TOWRE (Form A)<br>W-J Letter-Word ID (Form C) | CTOPP Color<br>CTOPP Letter<br>TOWRE (Form B)<br>W-J Letter-Word ID (Form A)<br>CTOPP Elision |
| K2_T1_BehavioralData_20200820.csv | K2_T2_BehavioralData_20200901.csv |

```{r load data, include=FALSE}

# t1
b1_data <- read_csv("K2_T1_BehavioralData_20200820.csv", col_names = TRUE, na = "N/A")

# t2
b2_data <- read_csv("K2_T2_BehavioralData_20200901.csv", col_names = TRUE, na = "N/A")

```

```{r clean data, include=FALSE}

# clean t1 data -----------------------------------------
b1_clean <- b1_data %>% 
  rename(pid = participant_id, 
         t1.age_year = k2_age,
         t1.age_months = k2_age_months,
         t1.age_days = k2_age_days,
         grade = k2_grade,
         t1.ctopp_c_raw = ctopp_color_raw,
         t1.ctopp_l_raw = ctopp_lttr_raw,
         t1.swe_raw = swe_raw2,
         t1.pde_raw = pde_raw_final,
         t1.swe_age = twre_scaled_age,
         t1.pde_age = pde_scaled_age,
         t1.swe_grade = swe_scaled_grade,
         t1.pde_grade = pde_scaled_grade,
         t1.wj_raw = wj_corr_final) %>% 
  select (pid, t1.age_year, t1.age_months, t1.age_days, grade,
          t1.ctopp_c_raw, t1.ctopp_l_raw,
          t1.swe_raw, t1.pde_raw, t1.swe_age, t1.pde_age,
          t1.swe_grade, t1.pde_grade, t1.wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(t1.age = (t1.age_year + (t1.age_months/12) + (t1.age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  select(-t1.age_months, -t1.age_days) %>% 
  select(pid, grade, t1.age, t1.age_year, everything())

# clean t2 data -----------------------------------------
b2_clean <- b2_data %>% 
  rename(pid = participant_id, 
         t2.age_year = k2_t2_b_age_years,
         t2.age_months = k2_t2_b_age_months,
         t2.age_days = k2_t2_b_age_days,
         grade = k2_t2_b_grade_year,
         t2.ctopp_c_raw = k2_t2_b_color_raw,
         t2.ctopp_l_raw = k2_t2_b_lttr_raw,
         t2.elision_raw = k2_t2_b_elision_raw,
         t2.swe_raw = k2_t2_b_swe_raw_fin,
         t2.pde_raw = k2_t2_b_pde_raw_fin,
         t2.swe_age = k2_t2_b_swe_agescal,
         t2.pde_age = k2_t2_b_pde_agescal,
         t2.swe_grade = k2_t2_b_swe_grdscal,
         t2.pde_grade = k2_t2_b_pde_grdscal,
         t2.wj_raw = k2_t2_b_wj_corr_nb) %>% 
  select (pid, t2.age_year, t2.age_months, t2.age_days, grade,
          t2.ctopp_c_raw, t2.ctopp_l_raw, t2.elision_raw,
          t2.swe_raw, t2.pde_raw, t2.swe_age, t2.pde_age,
          t2.swe_grade, t2.pde_grade, t2.wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(t2.age = (t2.age_year + (t2.age_months/12) + (t2.age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  select(-t2.age_months, -t2.age_days) %>% 
  select(pid, grade, t2.age, t2.age_year, everything())

# join time points--------------------------------------------

# all participants
b_all <- full_join(b1_clean, b2_clean, by = c("pid", "grade"))

# only participants who particpated in both time points
b_both <- inner_join(b1_clean, b2_clean, by = c("pid", "grade"))

```

## Participant breakdown
```{r grade breakdown, echo=FALSE, message=FALSE}

# make a table of grades
b1_grades <- b1_clean %>% group_by(grade) %>% summarise(T1 = n())
b2_grades <- b2_clean %>% group_by(grade) %>% summarise(T2 = n())
b_all_grades <- b_both %>% group_by(grade) %>% summarise(Both = n()) %>% 
  full_join(b1_grades, by="grade") %>% 
  full_join(b2_grades, by="grade") %>% 
  rename(Grade = grade) %>% 
  select(Grade, T1, T2, Both)

# adding a row for total numbers
b_all_grades <- rbind(b_all_grades, 
                      data.frame(Grade = 'Total', 
                                 T1 = sum(b_all_grades$T1),
                                 T2 = sum(b_all_grades$T2),
                                 Both = sum(b_all_grades$Both)))

# format table for rendered html
kable(b_all_grades) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                position = "left") %>% 
  row_spec(4, bold = T)

```

## Participants in both time points
```{r view combined data, echo=FALSE}

# interactive table with filters
datatable(b_both, rownames = FALSE, filter="top", options = list(pageLength = 10, scrollX = T))

```

# Initial Cleaning

For each time point: 

* Selected raw scores for CTOPP, TOWRE, W-J, plus scaled scores for TOWRE.
* Did not remove any subjects.

# Descriptive Statistics 

## T1
```{r t1 desc stats, echo=FALSE}

# t1 desc stats
b1_clean_desc <- b1_clean %>% 
  select(- c(pid, grade, t1.age_year)) %>%
  describe() 

# format table for rendered html
kable(b1_clean_desc, 
      digits = 3, # round to 3 digits
      caption = "T1 Descriptive Stats") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"),
                fixed_thead = T,
                font_size = 11)

```

## T2
```{r t2 desc stats, echo=FALSE}

# t2 desc stats
b2_clean_desc <- b2_clean %>% 
  select(- c(pid, grade, t2.age_year)) %>%
  describe()

# format table for rendered html
kable(b2_clean_desc, digits = 3, caption = "T2 Descriptive Stats") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                fixed_thead = T,
                font_size = 11)

```

```{r calculate means, echo=FALSE}

# calculate means by grade, use for plots later

# t1 means
b1_clean_means <- b1_clean %>%
  select(-pid, -t1.age_year) %>% 
  group_by(grade) %>% 
  summarise_all("mean", na.rm=TRUE)
# split by grade
b1_clean_means_K <- b1_clean_means %>% filter(grade == "K")
b1_clean_means_1 <- b1_clean_means %>% filter(grade == "1")
b1_clean_means_2 <- b1_clean_means %>% filter(grade == "2")

# t2 means
b2_clean_means <- b2_clean %>%
  select(-pid, -t2.age_year) %>% 
  group_by(grade) %>% 
  summarise_all("mean", na.rm=TRUE)
# split by grade
b2_clean_means_K <- b2_clean_means %>% filter(grade == "K")
b2_clean_means_1 <- b2_clean_means %>% filter(grade == "1")
b2_clean_means_2 <- b2_clean_means %>% filter(grade == "2")

```

# Visualizations

## CTOPP Rapid Color Naming

The Color Naming raw score is the **time (in seconds)** it takes to name all the colors on the page (lower raw score = better performance).

### Boxplots by grade

```{r ctopp color boxplots, echo=F, fig.width=10, fig.height=5}

# t1 boxplot by grade
p_cc_b1 <- b1_clean %>% 
  drop_na(t1.ctopp_c_raw) %>% 
  ggplot(aes(x = grade, y = t1.ctopp_c_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T1: CTOPP Color Raw Scores (N = 56)", 
       x = "Grade",
       y = "Raw Score (seconds)") +
  ylim(20, 80) +
  theme_light()

# t2 boxplot by grade
p_cc_b2 <- b2_clean %>% 
  drop_na(t2.ctopp_c_raw) %>% 
  ggplot(aes(x = grade, y = t2.ctopp_c_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T2: CTOPP Color Raw Scores (N = 52)", 
       x = "Grade",
       y = "Raw Score (seconds)") +
  ylim(20, 80) +
  theme_light()

# arrange side by side in rendered html
ggarrange(p_cc_b1, p_cc_b2, common.legend = TRUE, legend = "bottom")

```

Inspect high raw scores. Checking session notes:

T2

* BLC_264 (K) - Participant paused during assessment to ask a question.
* BLC_219 (2) - Difficult session, but no flags for CTOPP Color.
* BLC_226 (2) - No flags for CTOPP Color; participant was sitting far away from screen.

<mark>TO-DO: remove outliers?</mark>

```{r inspect and remove cc outliers, echo=F}

# look at t1 and t2 scores
outliers_cc <- b_all %>% 
  filter(pid == "BLC_264" | pid == "BLC_219" | pid == "BLC_226") %>% 
  select(pid, grade, t1.ctopp_c_raw, t2.ctopp_c_raw)

outliers_cc

```


### Density plots by grade

Dotted lines indicate raw score means by grade.

```{r ctopp color density by timept, echo=F, results='hide', fig.width=10, fig.height=5}

# t1 ctopp color density
p_cc_d1 <- b1_clean %>% 
  drop_na(t1.ctopp_c_raw) %>% 
  ggplot(aes(x = t1.ctopp_c_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b1_clean_means, 
             aes(xintercept = t1.ctopp_c_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T1: CTOPP Color Raw Scores",
       x = "Raw Score (seconds)",
       y = "Density") +
  theme_classic()

# t2 ctopp color density
p_cc_d2 <- b2_clean %>% 
  drop_na(t2.ctopp_c_raw) %>% 
  ggplot(aes(x = t2.ctopp_c_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = t2.ctopp_c_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T2: CTOPP Color Raw Scores",
       x = "Raw Score (seconds)",
       y = "Density") +
  theme_classic()

# arrange plots side by side
ggarrange(p_cc_d1, p_cc_d2, common.legend = TRUE, legend = "bottom")

```

### Scatterplots by age

```{r cc scatterplots, echo=F, message=F, fig.width=10, fig.height=5}

# t1 ctopp color scatter (x = age, y = raw)
p_cc_s1 <- b1_clean %>% 
  drop_na(t1.ctopp_c_raw) %>% 
  ggplot(aes(x = t1.age, y = t1.ctopp_c_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(20, 80) +
  labs(title = "T1: CTOPP Color Raw Scores (N = 56)", 
       x = "Age",
       y = "Raw Score (seconds)") +
  theme_light()

# t2 ctopp color scatter (x = age, y = raw)
p_cc_s2 <- b2_clean %>% 
  drop_na(t2.ctopp_c_raw) %>% 
  ggplot(aes(x = t2.age, y = t2.ctopp_c_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(20, 80) +
  labs(title = "T2: CTOPP Color Raw Scores (N = 52)", 
       x = "Age",
      y = "Raw Score (seconds)") +
  theme_light()

# arrange plots side by side
ggarrange(p_cc_s1, p_cc_s2, common.legend = TRUE, legend = "bottom")

```

```{r cc linear regression with age}

# t1 ctopp color (raw ~ age)
lm_cc_t1 <- lm(t1.ctopp_c_raw ~ t1.age, data = b1_clean)
print(summary(lm_cc_t1))

# t2 ctopp color (raw ~ age)
lm_cc_t2 <- lm(t2.ctopp_c_raw ~ t2.age, data = b2_clean)
print(summary(lm_cc_t2))

```

### T1-T2

Subjects who completed both time points.

```{r ctopp color t1-t2, echo=F, message=F, fig.width=8, fig.height=5}

# t1-t2 ctopp color scatterplot of raw scores
p_cc_12 <- ggplot(b_both, aes(x = t1.ctopp_c_raw, y = t2.ctopp_c_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") + 
  labs(title = "T1-T2: CTOPP Color Raw Scores (N = 43)", 
       x = "T1 Raw Score",
       y = "T2 Raw Score") +
  theme_light()

p_cc_12

# linear regression
lm_cc <- lm(t2.ctopp_c_raw ~ t1.ctopp_c_raw, data = b_both)
print(summary(lm_cc))

```

## CTOPP Rapid Letter Naming

The Letter Naming raw score is the **time (in seconds)** it takes to name all the letters on the page (lower raw score = better performance).

### Boxplots by grade
```{r ctopp letter boxplots, echo=F, fig.width=10, fig.height=5}

# t1 ctopp letter boxplot
p_cl_b1 <- b1_clean %>% 
  drop_na(t1.ctopp_l_raw) %>% 
  ggplot(aes(x = grade, y = t1.ctopp_l_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T1: CTOPP Letter Raw Scores (N = 55)", 
       x = "Grade",
       y = "Raw Score (seconds)") +
  ylim(10, 80) +
  theme_light()

# t2 ctopp letter boxplot
p_cl_b2 <- b2_clean %>% 
  drop_na(t2.ctopp_l_raw) %>% 
  ggplot(aes(x = grade, y = t2.ctopp_l_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T2: CTOPP Letter Raw Scores (N = 53)", 
       x = "Grade",
       y = "Raw Score (seconds)") +
  ylim(10, 80) +
  theme_light()

# arrange side by side
ggarrange(p_cl_b1, p_cl_b2, common.legend = TRUE, legend = "bottom")

```

Inspect high raw scores. Checking session notes:

T1 

* BLC_247 (1) - No flags for CTOPP Letter.

T2

* BLC_233 (K) - Struggled a little with the letter "c" (had to take longer to think).
* BLC_258 (1) - No flags for CTOPP Letter.

<mark>TO-DO: remove outliers?</mark>

```{r inspect and remove cl outliers, echo=F}

# look at t1 and t2 scores
outliers_cl <- b_all %>% 
  filter(pid == "BLC_247" | pid == "BLC_258" | pid == "BLC_233") %>% 
  select(pid, grade, t1.ctopp_l_raw, t2.ctopp_l_raw)

outliers_cl

```

### Density plots by grade

Dotted lines indicate raw score means by grade.

```{r ctopp letter density by timept, echo=F, results='hide', fig.width=10, fig.height=5}

# t1 ctopp letter density
p_cl_d1 <- b1_clean %>% 
  drop_na(t1.ctopp_l_raw) %>% 
  ggplot(aes(x = t1.ctopp_l_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b1_clean_means, 
             aes(xintercept = t1.ctopp_l_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T1: CTOPP Letter Raw Scores (N = 55)",
       x = "Raw Score (seconds)",
       y = "Density") +
  theme_classic()

# t2 ctopp letter density
p_cl_d2 <- b2_clean %>% 
  drop_na(t2.ctopp_l_raw) %>% 
  ggplot(aes(x = t2.ctopp_l_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = t2.ctopp_l_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T2: CTOPP Letter Raw Scores (N = 53)",
       x = "Raw Score (seconds)",
       y = "Density") +
  theme_classic()

# arrange side by side
ggarrange(p_cl_d1, p_cl_d2, common.legend = TRUE, legend = "bottom")

```

### Scatter plots by age
```{r ctopp letter scatterplots, echo=F, message=F, fig.width=10, fig.height=5}

# t1 ctopp letter scatter (x = age, y = raw)
p_cl_s1 <- b1_clean %>% 
  drop_na(t1.ctopp_l_raw) %>% 
  ggplot(aes(x = t1.age, y = t1.ctopp_l_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 80) +
  labs(title = "T1: CTOPP Letter Raw Scores (N = 55", 
       x = "Age", 
       y = "Raw Score (seconds)") +
  theme_light()

# t2 ctopp letter scatter (x = age, y = raw)
p_cl_s2 <- b2_clean %>% 
  drop_na(t2.ctopp_l_raw) %>% 
  ggplot(aes(x = t2.age, y = t2.ctopp_l_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 80) +
  labs(title = "T2: CTOPP Letter Raw Scores (N = 53)",         
       x = "Age",
       y = "Raw Score (seconds)") +
  theme_light()

# arrange side by side
ggarrange(p_cl_s1, p_cl_s2, common.legend = TRUE, legend = "bottom")

```

### T1-T2

Subjects who completed both time points.

```{r ctopp letter t1-t2, echo=F, message=F, fig.width=8, fig.height=5}

# t1-t2 ctopp letter scatterplot of raw scores
p_cl_12 <- b_both %>% 
  drop_na(t1.ctopp_l_raw, t2.ctopp_l_raw) %>% 
  ggplot(aes(x = t1.ctopp_l_raw, y = t2.ctopp_l_raw)) + 
    geom_point(aes(color = grade)) +
    geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") + 
    labs(title = "T1-T2: CTOPP Letter Raw Scores (N = 42)", 
         x = "T1 Raw Score",
         y = "T2 Raw Score") +
  theme_light()

p_cl_12

# linear regression (t2 raw ~ t1 raw)
lm_cl <- lm(t2.ctopp_l_raw ~ t1.ctopp_l_raw, data = b_both)
print(summary(lm_cl))

```

## CTOPP Elision (T2 Only)

The Elision raw score is the **number of items answered correctly** before reaching the ceiling rule (3 in a row incorrect).

```{r elision boxplot, echo=F, fig.width=10, fig.height=5}

# t2 elision boxplot
p_ce_b2 <- b2_clean %>% 
  drop_na(t2.elision_raw) %>% 
  ggplot(aes(x = grade, y = t2.elision_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T2: CTOPP Elision Raw Scores (N = 53)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  theme_light()

p_ce_b2

```

Inspect low scores. Checking session notes:

* BLC_245 (2) - Participant wanted the test to be over; was asking "how many more?"

```{r elision plots, echo=F, message=F, fig.width=10, fig.height=5}

# t2 ctopp elision density
p_ce_d2 <- b2_clean %>% 
  ggplot(aes(x = t2.elision_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = t2.elision_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T2: CTOPP Elision Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# t2 ctopp elision scatter (x = age, y = raw)
p_ce_s2 <- b2_clean %>% 
  ggplot(aes(x = t2.age, y = t2.elision_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) + 
  labs(title = "T2: CTOPP Elision Raw Scores (N = 53)", 
       x = "Age",
       y = "Raw Score (# correct)") +
  theme_classic()

# arrange side by side
ggarrange(p_ce_d2, p_ce_s2, common.legend = TRUE, legend = "bottom")

```

## TOWRE Sight Word Efficiency

The Sight Word Efficiency (SWE) raw score is the **number of words read correctly** in 45 seconds.

### Boxplots by grade
```{r swe boxplots, echo=F, fig.width=10, fig.height=5}

# t1 swe boxplot
p_swe_b1 <- b1_clean %>% 
  drop_na(t1.swe_raw) %>% 
  ggplot(aes(x = grade, y = t1.swe_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T1: SWE Raw Scores (N = 56)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  ylim(0, 75) +
  theme_light()

# t2 swe boxplot
p_swe_b2 <- b2_clean %>% 
  drop_na(t2.swe_raw) %>% 
  ggplot(aes(x = grade, y = t2.swe_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T2: SWE Raw Scores (N = 53)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  ylim(0, 75) +
  theme_light()

# arrange side by side
ggarrange(p_swe_b1, p_swe_b2, common.legend = TRUE, legend = "bottom")

```

Inspect low scores. Checking session notes:

T1

* BLC_205 (2) - No flags for SWE test.

T2

* BLC_205 (2) - No flags for SWE test.
* BLC_258 (1) - No flags for SWE test.

Both BLC_205 and BLC_258 show lower scores for TOWRE as a whole.

<mark>TO-DO: remove outliers?</mark>

```{r inspect and remove swe outliers, echo=F}

# look at t1 and t2 scores
outliers_swe <- b_all %>% 
  filter(pid == "BLC_205" | pid == "BLC_258") %>% 
  select(pid, grade, t1.swe_raw, t2.swe_raw, t1.pde_raw, t2.pde_raw)

outliers_swe

```

### Density plots by grade
```{r swe density, echo=F, fig.width=10, fig.height=5}

# t1 swe density
p_swe_d1 <- b1_clean %>% 
  drop_na(t1.swe_raw) %>% 
  ggplot(aes(x = t1.swe_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b1_clean_means, 
             aes(xintercept = t1.swe_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T1: SWE Raw Scores (N = 56)",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# t2 swe density
p_swe_d2 <- b2_clean %>% 
  drop_na(t2.swe_raw) %>% 
  ggplot(aes(x = t2.swe_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = t2.swe_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T2: SWE Raw Scores (N = 53)",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# arrange side by side
ggarrange(p_swe_d1, p_swe_d2, common.legend = TRUE, legend = "bottom")

```

### Scatterplots by age
```{r swe scatter, echo=F, message=F, fig.width=10, fig.height=5}

# t1 swe scatter (x = age, y = raw)
p_swe_s1 <- b1_clean %>% 
  drop_na(t1.swe_raw) %>% 
  ggplot(aes(x = t1.age, y = t1.swe_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = F, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 75) +
  labs(title = "T1: SWE Raw Scores (N = 56)", 
       x = "Age", 
       y = "Raw Score (# correct)") +
  theme_light()

# t2 swe scatter (x = age, y = raw)
p_swe_s2 <- b2_clean %>% 
  drop_na(t2.swe_raw) %>% 
  ggplot(aes(x = t2.age, y = t2.swe_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = F, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 75) +
  labs(title = "T2: SWE Raw Scores (N = 53)",         
       x = "Age",
       y = "Raw Score (# correct)") +
  theme_light()

# arrange side by side
ggarrange(p_swe_s1, p_swe_s2, common.legend = TRUE, legend = "bottom")

```

### T1-T2

```{r swe t1-t2, echo=F, message=F, fig.width=8, fig.height=5}

# t1-t2 swe scatterplot of raw scores
p_swe_12 <- b_both %>% 
  drop_na(t1.swe_raw, t2.swe_raw) %>% 
  ggplot(aes(x = t1.swe_raw, y = t2.swe_raw)) + 
    geom_point(aes(color = grade)) +
    geom_smooth(method = "lm", se = F, linetype = "solid", color = "grey") + 
    labs(title = "T1-T2: SWE Raw Scores (N = 42)", 
         x = "T1 Raw Score",
         y = "T2 Raw Score") +
  theme_light()

p_swe_12

# linear regression (t2 raw ~ t1 raw)
lm_swe <- lm(t2.swe_raw ~ t1.swe_raw, data = b_both)
print(summary(lm_swe))

```

## TOWRE Phonemic Decoding Efficiency

The Phonemic Decoding Efficiency (PDE) raw score is the **number of non-words read correctly** in 45 seconds.

### Boxplots by grade
```{r pde boxplots, echo=F, fig.width=10, fig.height=5}

# t1 pde boxplot
p_pde_b1 <- b1_clean %>% 
  drop_na(t1.pde_raw) %>% 
  ggplot(aes(x = grade, y = t1.pde_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T1: PDE Raw Scores (N = 55)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  ylim(0, 50) +
  theme_light()

# t2 pde boxplot
p_pde_b2 <- b2_clean %>% 
  drop_na(t2.pde_raw) %>% 
  ggplot(aes(x = grade, y = t2.pde_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T2: PDE Raw Scores (N = 50)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  ylim(0, 50) +
  theme_light()

# arrange side by side
ggarrange(p_pde_b1, p_pde_b2, common.legend = TRUE, legend = "bottom")

```

### Density plots by grade
```{r pde density, echo=F, fig.width=10, fig.height=5}

# t1 pde density
p_pde_d1 <- b1_clean %>% 
  drop_na(t1.pde_raw) %>% 
  ggplot(aes(x = t1.pde_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b1_clean_means, 
             aes(xintercept = t1.pde_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T1: PDE Raw Scores (N = 55)",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# t2 pde density
p_pde_d2 <- b2_clean %>% 
  drop_na(t2.pde_raw) %>% 
  ggplot(aes(x = t2.pde_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = t2.pde_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T2: PDE Raw Scores (N = 50)",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# arrange side by side
ggarrange(p_pde_d1, p_pde_d2, common.legend = TRUE, legend = "bottom")

```

### Scatterplots by age
```{r pde scatter, echo=F, message=F, fig.width=10, fig.height=5}

# t1 pde scatter (x = age, y = raw)
p_pde_s1 <- b1_clean %>% 
  drop_na(t1.pde_raw) %>% 
  ggplot(aes(x = t1.age, y = t1.pde_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = F, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 50) +
  labs(title = "T1: PDE Raw Scores (N = 55)", 
       x = "Age", 
       y = "Raw Score (# correct)") +
  theme_light()

# t2 pde scatter (x = age, y = raw)
p_pde_s2 <- b2_clean %>% 
  drop_na(t2.pde_raw) %>% 
  ggplot(aes(x = t2.age, y = t2.pde_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = F, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 50) +
  labs(title = "T2: PDE Raw Scores (N = 50)",         
       x = "Age",
       y = "Raw Score (# correct)") +
  theme_light()

# arrange side by side
ggarrange(p_pde_s1, p_pde_s2, common.legend = TRUE, legend = "bottom")

```

### T1-T2
```{r pde t1-t2, echo=F, message=F, fig.width=8, fig.height=5}

# t1-t2 pde scatterplot of raw scores
p_pde_12 <- b_both %>% 
  drop_na(t1.pde_raw, t2.pde_raw) %>% 
  ggplot(aes(x = t1.pde_raw, y = t2.pde_raw)) + 
    geom_point(aes(color = grade)) +
    geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") + 
    labs(title = "T1-T2: PDE Raw Scores (N = 38)", 
         x = "T1 Raw Score",
         y = "T2 Raw Score") +
  theme_light()

p_pde_12

# linear regression (t2 raw ~ t1 raw)
lm_pde <- lm(t2.pde_raw ~ t1.pde_raw, data = b_both)
print(summary(lm_pde))

```

## TOWRE Grade-based Scaled Scores
```{r towre t1-t2 difference, echo=F}

# TO-DO: Magnitude of difference between T1 and T2 grade-based scaled scores

b_both_towre_grscal <- b_both %>% 
  select(pid, grade, 
         t1.swe_grade, t2.swe_grade, 
         t1.pde_grade, t2.pde_grade) %>% 
  filter(grade != 'K') %>% 
  mutate(swe_difference = t2.swe_grade - t1.swe_grade) %>% 
  mutate(pde_difference = t2.pde_grade - t1.pde_grade)

```

## W-J Letter-Word Identification

The W-J raw score is the **number of items answered correctly** before reaching the ceiling rule (6 highest items that were administered are incorrect).

### Boxplots by grade
```{r w-j boxplots, echo=F, fig.width=10, fig.height=5}

# t1 w-j boxplot
p_wj_b1 <- b1_clean %>% 
  drop_na(t1.wj_raw) %>% 
  ggplot(aes(x = grade, y = t1.wj_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T1: W-J Raw Scores (N = 57)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  ylim(0, 70) +
  theme_light()

# t2 w-j boxplot
p_wj_b2 <- b2_clean %>% 
  drop_na(t2.wj_raw) %>% 
  ggplot(aes(x = grade, y = t2.wj_raw)) +
  geom_boxplot(aes(color = grade)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, binwidth=1) +
  labs(title = "T2: W-J Raw Scores (N = 53)", 
       x = "Grade",
       y = "Raw Score (# correct)") +
  ylim(0, 70) +
  theme_light()

# arrange side by side
ggarrange(p_wj_b1, p_wj_b2, common.legend = TRUE, legend = "bottom")

```

Inspect low scores. Checking session notes:

T2

* BLC_258 (1) - Flagged for audio lag/distortion during this test; participant also has low scores across the board.

### Density plots by grade
```{r w-j density, echo=F, fig.width=10, fig.height=5}

# t1 wj density
p_wj_d1 <- b1_clean %>% 
  drop_na(t1.wj_raw) %>% 
  ggplot(aes(x = t1.wj_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b1_clean_means, 
             aes(xintercept = t1.wj_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T1: W-J Raw Scores (N = 57)",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# t2 wj density
p_wj_d2 <- b2_clean %>% 
  drop_na(t2.wj_raw) %>% 
  ggplot(aes(x = t2.wj_raw, fill = grade)) +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = t2.wj_raw, color = grade),
             linetype = "dashed") +
  labs(title = "T2: W-J Raw Scores (N = 53)",
       x = "Raw Score (# correct)",
       y = "Density") +
  theme_classic()

# arrange side by side
ggarrange(p_wj_d1, p_wj_d2, common.legend = TRUE, legend = "bottom")

```

### Scatterplots by age
```{r w-j scatter, echo=F, message=F, fig.width=10, fig.height=5}

# t1 w-J scatter (x = age, y = raw)
p_wj_s1 <- b1_clean %>% 
  drop_na(t1.wj_raw) %>% 
  ggplot(aes(x = t1.age, y = t1.wj_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 70) +
  labs(title = "T1: W-J Raw Scores (N = 57)", 
       x = "Age", 
       y = "Raw Score (# correct)") +
  theme_light()

# t2 w-j scatter (x = age, y = raw)
p_wj_s2 <- b2_clean %>% 
  drop_na(t2.wj_raw) %>% 
  ggplot(aes(x = t2.age, y = t2.wj_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") +
  scale_x_continuous(breaks = seq(5, 9, by = 0.5)) +
  ylim(0, 70) +
  labs(title = "T2: W-J Raw Scores (N = 53)",         
       x = "Age",
       y = "Raw Score (# correct)") +
  theme_light()

# arrange side by side
ggarrange(p_wj_s1, p_wj_s2, common.legend = TRUE, legend = "bottom")

```

### T1-T2
```{r w-j t1-t2, echo=FALSE, message=F, fig.width=8, fig.height=6}

# t1-t2 w-j scatterplot of raw scores
p_wj_12 <- b_both %>% 
  drop_na(t1.wj_raw, t2.wj_raw) %>% 
  ggplot(aes(x = t1.wj_raw, y = t2.wj_raw)) + 
    geom_point(aes(color = grade)) +
    geom_smooth(method = "lm", se = T, linetype = "solid", color = "grey") + 
    labs(title = "T1-T2: W-J Raw Scores (N = 43)", 
         x = "T1 Raw Score",
         y = "T2 Raw Score") +
  theme_light()

p_wj_12

# linear regression (t2 raw ~ t1 raw)
lm_wj <- lm(t2.wj_raw ~ t1.wj_raw, data = b_both)
print(summary(lm_wj))

```