---
title: 'K-2 T2 Behavioral Data (RA Version: Exploration in R)'
author: "Cha Cha Pillai, Mira Guleri, Angie Wang"
date: "8/27/2020"
output:
  pdf_document: default
  pdf: default
---

# About this File

This is the RA practice version! See *K2_T1_T2_Behavioral.Rmd* for the actual T1-T2 script.

# Markdown
Text in these white spaces will appear in your rendered file as text.
<!--But this text won't appear!-->

# This creates a heading
## This is a smaller heading
### An even smaller heading!

Click the Knit button to render the file and see what the text looks like.        

```{r set up, include=FALSE}
# This is our first code chunk. In the curly braces / chunk header above, 
# 'r' (which is automatically inserted) tells us that this is an R code chunk.
# We then named this chunk 'set up,' and specified that the contents of the chunk 
# shouldn't appear in the rendered file with the 'include = FALSE' argument.

# First, we load the packages that we'll be using. The packages have to 
# be installed before you load them, which you can do using install.packages().

library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(pastecs)    # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arranging plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling

# You'll also want to check what your working directory is, so R knows where 
# to look for files. Try getwd() and setwd(). Alternatively, you can go to the
# top menu bar on your computer > Session > Set Working Directory > Choose Directory.

```

\newpage
# Background and Participants

* Grades: K, 1, 2
* Ages: 5 to 8

## K-2 T2

Second time point collected remotely via Zoom, May-June 2020 (end of the school year)

* Assessments: CTOPP Color, CTOPP Letter, CTOPP Elision, TOWRE, W-J LWID
* Data file: K2_T2_BehavioralData_20200728.csv
* N = 57

## K-2 T1

First time point collected in-person, October 2019 (beginning of the school year)

* Assessments: CTOPP Color, CTOPP Letter, TOWRE, W-J LWID
* Data file: K2_T1_BehavioralData_20200820.csv
* N = 54


```{r clean t2 data, include=FALSE}
# Next, we read in the .csv file. We included additional arguments in read_csv():
# (1) Set col_names as TRUE so the first row of the input will be used as the 
# column names, and will not be included in the data frame. (2) Specify that the 
# string "N/A" in the data should be interpreted as a missing value.

# read in T2 data
b2_data <- read_csv("K2_T2_BehavioralData_20200728.csv", 
                    col_names = TRUE, 
                    na = "N/A")

# look at the class of every column in the dataframe
#sapply(b2_data, class)

# you can do something like this to change the class of a specific column
#b2_data$k2_t2_b_grade_year <- as.factor(b2_data$k2_t2_b_grade_year)

# view your dataframe
#View(b2_data)

# After checking that the data looks normal, we use pipes (%>%) for a series 
# of actions on our original data frame, b2_data. The cleaned result is 
# assigned to a new variable, b2_clean.

# look at existing column names
#colnames(b2_data)

# clean the data
b2_clean <- b2_data %>% 
  # rename columns (new_name = old_name)
  rename(pid = participant_id, 
         age_year = k2_t2_b_age_years,
         age_months = k2_t2_b_age_months,
         age_days = k2_t2_b_age_days,
         grade = k2_t2_b_grade_year,
         ctopp_c_raw = k2_t2_b_color_raw,
         ctopp_l_raw = k2_t2_b_lttr_raw,
         elision_raw = k2_t2_b_elision_raw,
         swe_raw = k2_t2_b_swe_raw_fin,
         pde_raw = k2_t2_b_pde_raw_fin,
         swe_age = k2_t2_b_swe_agescal,
         pde_age = k2_t2_b_pde_agescal,
         swe_grade = k2_t2_b_swe_grdscal,
         pde_grade = k2_t2_b_pde_grdscal,
         wj_raw = k2_t2_b_wj_corr_nb) %>% 
  # select columns we want
  select (pid, age_year, age_months, age_days, grade,
          ctopp_c_raw, ctopp_l_raw, elision_raw,
          swe_raw, pde_raw, swe_age, pde_age,
          swe_grade, pde_grade, wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(age = (age_year + (age_months/12) + (age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  # de-select columns that we don't need anymore
  select(-age_months, -age_days) %>% 
  # reorder columns
  select(pid, grade, age, age_year, everything())

```

```{r clean t1 data, include=FALSE}

b1_data <- read_csv("K2_T1_BehavioralData_20200820.csv", 
                    col_names = TRUE, 
                    na = "N/A")

b1_clean <- b1_data %>% 
  rename(pid = participant_id, 
         t1.age_year = k2_age,
         age_months = k2_age_months,
         age_days = k2_age_days,
         grade = k2_grade,
         t1.ctopp_c_raw = ctopp_color_raw,
         t1.ctopp_l_raw = ctopp_lttr_raw,
         t1.swe_raw = swe_raw2,
         t1.pde_raw = pde_raw_final,
         t1.swe_age = twre_scaled_age,
         t1.pde_age = pde_scaled_age,
         t1.swe_grade = swe_scaled_grade,
         t1.pde_grade = pde_scaled_grade,
         t1.wj_raw = wj_corr_final) %>% 
  select (pid, t1.age_year, age_months, age_days, grade,
          t1.ctopp_c_raw, t1.ctopp_l_raw,
          t1.swe_raw, t1.pde_raw, t1.swe_age, t1.pde_age,
          t1.swe_grade, t1.pde_grade, t1.wj_raw) %>% 
  # create new column for age as a decimal value
  mutate(t1.age = (t1.age_year + (age_months/12) + (age_days/365))) %>% 
  # reorder grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,"K", "1", "2")) %>% 
  # de-select columns that we don't need anymore
  select(-age_months, -age_days) %>% 
  # reorder columns
  select(pid, grade, t1.age, t1.age_year, everything())

```

# Data Cleaning

For each time point: 

* Selected raw scores for CTOPP, TOWRE, W-J, plus scaled scores for TOWRE.
* Did not remove any subjects.

Then combined time points.

* N = 43 participants who participated in both time points.

```{r combine time points, echo=FALSE, message=FALSE}

# join cleaned data from both time points
b_both <- inner_join(b1_clean, b2_clean, by = c("pid", "grade"))

# make a table of grades
b1_grades <- b1_clean %>% group_by(grade) %>% summarise(T1 = n())
b2_grades <- b2_clean %>% group_by(grade) %>% summarise(T2 = n())

# grades all together
b_all_grades <- b_both %>% group_by(grade) %>% summarise(Both = n()) %>% 
  full_join(b1_grades, by="grade") %>% 
  full_join(b2_grades, by="grade") %>% 
  rename(Grade = grade) %>% 
  select(Grade, T1, T2, Both)

# adding totals to the grade table
b_all_grades <- rbind(b_all_grades, 
                      data.frame(Grade = 'Total', 
                                 T1 = sum(b_all_grades$T1),
                                 T2 = sum(b_all_grades$T2),
                                 Both = sum(b_all_grades$Both)))

b_all_grades

# format table for rendered file
kable(b_all_grades, 
      format = "latex", 
      caption = "Grade breakdown",
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>% 
  row_spec(4, bold = TRUE)

```


\newpage <!--putting this on a new page in the rendered file-->
# Extra Practice: Manipulating dataframes with dplyr verbs 
Here's a useful dplyr cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf

### Exercise 1
Create a new data frame with just the participant id, age, grade, and CTOPP Elision data. Try reordering the columns, like putting grade ahead of everything else.

```{r dplyr extra practice #1}

# 1. Solution


```

### Exercise 2a
Create a new data frame with just 2nd grade W-J data. (Hint: Use `filter()`). You should end up with 19 subjects.

### Exercise 2b
Still working with the same data frame in 2a (Hint: Use piping!), pick out subjects who scored 60 or above. You should end up with 8 subjects.

```{r dplyr extra practice #2}
# 2a. Solution



# 2b. Solution



```

### Exercise 3
Create a new dataframe with 1st and 2nd grade data. (Hint: Use logical operators to get both grade levels)

```{r dplyr extra practice #3}
# 3. Solution


```

### Exercise 4a
The `arrange()` function orders rows by values of a column or columns (low to high). Create a new dataframe where you arrange the `b2_clean` data by age (low to high)

### Exercise 4b
Use `drop_na()` to drop rows (subjects) that have any NA values in their scores.

```{r dplyr extra practice #4}
# 4a. Solution


# 4b. Solution


```

\newpage 
# Descriptive Statistics 
```{r descriptive stats, echo=FALSE}

# descriptive stats
b2_clean_descr <- b2_clean %>%
  select(- pid, - grade, - age_year) %>% # remove unnecessary columns
  stat.desc() # using stat.desc() from pastecs package

# comment this section out if it's currently throwing an error (this is just
# formatting stuff for now):
# put the desc stats in a nice table in our rendered file
kable(b2_clean_descr, 
      digits = 3, # round to 3 digits
      format = "latex", 
      caption = "K-2 T2 Descriptive Statistics",
      booktabs = TRUE) %>%
  # more formatting, from the kableExtra package
  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) 


# Click the "Knit" button to render the file and view the table of
# descriptive statistics


# Calculate means by grade, use for plots later
b2_clean_means <- b2_clean %>%
  select(-pid, -age_year) %>% 
  group_by(grade) %>% 
  summarise_all("mean", na.rm=TRUE)

# TO-DO: Put means by grade in table for rendered file

b2_clean_means_K <- b2_clean_means %>% filter(grade == "K")
b2_clean_means_1 <- b2_clean_means %>% filter(grade == "1")
b2_clean_means_2 <- b2_clean_means %>% filter(grade == "2")

```

\newpage
# Visualizations

We are using the ggplot2 package (part of the tidyverse) to visualize the data.

* Learn more about the package here: https://ggplot2.tidyverse.org/.

* Cheat Sheet: https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf

In most cases you will:

1. Start with `ggplot()`
2. Supply a dataset in the first argument of `ggplot()`
3. Supply aesthetic mapping with `aes()`
4. Add on a `geom` function (like `geom_point()` or `geom_histogram()`)
5. Add on other layers and specifications (like editing the names of your axes, color, etc.)

## Examples using CTOPP Elision data

The Elision raw score is the number of items answered correctly before reaching the ceiling rule (3 in a row incorrect).

Basic scatterplot of elision raw scores:

```{r ggplot elision example basic, echo=FALSE}

# 1. We start with ggplot(). 

# 2. We specify the dataset to use in the graph in the first argument
# of ggplot(), in this case, b2_elision.

# 3. The x and y arguments of aes() specify which variables to map to the 
# x and y axes. Here, we specify that we want age on the x-axis and elision 
# raw scores on the y-axis.

# 4. We add on the function geom_point() to add a layer of points to a plot

ggplot(b2_clean, aes(x = age, y = elision_raw)) + 
  geom_point() # scatterplot

# We get a basic scatterplot!

```

\newpage
To enhance our scatterplot, we add on more layers and specifications:

```{r ggplot elision example enhanced, echo=FALSE}

# Adding color = grade as an argument in aes() to color-code the points by grade-level
p_ce_1 <- ggplot(b2_clean, aes(x = age, y = elision_raw, color = grade)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) + # adjusting x axis
  labs(title = "CTOPP Elision Raw Scores", # adding titles and labels with labs()
       x = "Age",
       y = "Raw Score (# correct)",
       caption = "This is my caption.") +
  theme_classic() # changing overall theme

p_ce_1

```


```{r ggplot elision other plots, echo=FALSE, message=FALSE}

# Make a CTOPP Elision Density Plot - plot densities of elision raw data by grade

p_ce_2 <- ggplot(b2_clean, aes(x = elision_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = elision_raw, color = grade),
             linetype = "dashed") +
  labs(title = "CTOPP Elision Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_ce_2


# Make a CTOPP Elision Histogram (of all scores)

p_ce_3 <- ggplot(b2_clean, aes(x = elision_raw)) +
  geom_histogram(fill = "lightblue") +
  labs(title = "CTOPP Elision Raw Scores",
       x = "Raw Score (# correct)",
       y = "Count") +
  theme_classic()

p_ce_3


# Make a CTOPP Elision Box Plot - by grade

p_ce_4 <- ggplot(b2_clean, aes(grade, elision_raw), fill = grade) +
  geom_boxplot(aes(fill = grade), alpha = 0.3) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "CTOPP Elision Raw Scores",
       x = "Grade",
       y = "Raw Score (# Correct)") +
  theme_classic()

p_ce_4

```

\newpage
## CTOPP Rapid Naming

### Rapid Color Naming

The Color Naming raw score is the time (in seconds) it takes to name all the colors on the page.

```{r ctopp color naming plots, echo=FALSE}

# TO-DO: Scatter (x = Age and y = Color Raw)

p_cc_1 <- ggplot(b2_clean, aes(x = age, y = ctopp_c_raw, color = grade)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) + # adjusting x axis
  labs(title = "CTOPP Color Raw Scores", # adding titles and labels with labs()
       x = "Age",
       y = "Raw Score (time to finish in seconds)",
       caption = "Lower raw score = better performance; Higher raw = worse performance.") +
  theme_classic() # changing overall theme

p_cc_1

# TO-DO: Density (Color Raw Scores)

p_cc_2 <- ggplot(b2_clean, aes(x = ctopp_c_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = ctopp_c_raw, color = grade),
             linetype = "dashed") +
  labs(title = "CTOPP Color Raw Scores",
       x = "Raw Score (time to finish in seconds)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_cc_2

```

\newpage
### Rapid Letter Naming

The Letter Naming raw score is the time (in seconds) it takes to name all the letters on the page.

```{r ctopp letter naming plots, echo=FALSE}

# TO-DO: Scatter (x = Age and y = Letter Raw)

p_cl_1 <- ggplot(b2_clean, aes(x = age, y = ctopp_l_raw, color = grade)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) +
  labs(title = "CTOPP Letter Raw Scores", 
       x = "Age",
       y = "Raw Score (time to finish in seconds)",
       caption = "Lower raw score = better performance; Higher raw = worse performance.") +
  theme_classic()

p_cl_1

# TO-DO: Density (Letter Raw Scores)

p_cl_2 <- ggplot(b2_clean, aes(x = ctopp_l_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = ctopp_l_raw, color = grade),
             linetype = "dashed") +
  labs(title = "CTOPP Letter Raw Scores",
       x = "Raw Score (time to finish in seconds)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_cl_2

```

\newpage
### CTOPP Comparisons

```{r ctopp comparisons, echo=FALSE, message=FALSE, warning=FALSE}

# TO-DO: Scatter (Color and Letter)
p_cc_cl <- ggplot(b2_clean, aes(x = ctopp_c_raw, y = ctopp_l_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "CTOPP Color and Letter (raw scores)", 
       x = "Color Raw Score (time to finish in seconds)",
       y = "Letter Raw Score (time to finish in seconds)") +
  theme_classic()

p_cc_cl

# TO-DO: Scatter (Letter and Elision)
p_cl_ce <- ggplot(b2_clean, aes(x = ctopp_l_raw, y = elision_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "CTOPP Letter and Elision (raw scores)", 
       x = "Letter Raw Score (time to finish in seconds)",
       y = "Elision Raw Score (# correct)") +
  theme_classic()

p_cl_ce

# TO-DO: Scatter (Color and Elision)
p_cc_ce <- ggplot(b2_clean, aes(x = ctopp_c_raw, y = elision_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "CTOPP Color and Elision (raw scores)", 
       x = "Color Raw Score (time to finish in seconds)",
       y = "Elision Raw Score (# correct)") +
  theme_classic()

p_cc_ce
  
```

\newpage
## TOWRE

### Sight Word Efficiency (SWE)

The SWE raw score is the number of words read correctly in 45 seconds.

```{r towre-swe plots, echo=FALSE}

# TO-DO: Scatter (x = Age and y = Raw)

p_swe_1 <- ggplot(b2_clean, aes(x = age, y = swe_raw, color = grade)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) +
  labs(title = "Sight Word Efficiency Raw Scores", 
       x = "Age",
       y = "Raw Score (# correct)") +
  theme_classic()

p_swe_1

# TO-DO: Density (Raw Scores by grade)

p_swe_2 <- ggplot(b2_clean, aes(x = swe_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = swe_raw, color = grade),
             linetype = "dashed") +
  labs(title = "Sight Word Efficiency Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_swe_2


```

\newpage
### Phonemic Decoding Efficiency (PDE)

The PDE raw score is the number of words read correctly in 45 seconds.

```{r towre-pde plots, echo=FALSE}

# TO-DO: Scatter (x = Age and y = Raw)

p_pde_1 <- ggplot(b2_clean, aes(x = age, y = pde_raw, color = grade)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) +
  labs(title = "Phonemic Decoding Efficiency Raw Scores", 
       x = "Age",
       y = "Raw Score (# correct)") +
  theme_classic()

p_pde_1

# TO-DO: Density (Raw Scores by grade)

p_pde_2 <- ggplot(b2_clean, aes(x = pde_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = pde_raw, color = grade),
             linetype = "dashed") +
  labs(title = "Phonemic Decoding Efficiency Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_pde_2


```

### TOWRE Comparisons
```{r towre comparisons, echo=FALSE, message=FALSE, warning=FALSE}

# Scatter (SWE and PDE Raw)

p_swe_pde_1 <- ggplot(b2_clean, aes(x = swe_raw, y = pde_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "TOWRE SWE and PDE (raw scores)", 
       x = "SWE Raw Score (words read correctly)",
       y = "PDE Raw Score (nonwords read correctly)",
       caption = "excludes 3 subjects with NAs") +
  theme_classic()

p_swe_pde_1

# Scatter (SWE and PDE Raw) with regression lines by grade
p_swe_pde_2 <- ggplot(b2_clean, aes(x = swe_raw, y = pde_raw, color = grade)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, linetype = "solid") + 
  labs(title = "TOWRE SWE and PDE (raw scores)", 
       x = "SWE Raw Score (words read correctly)",
       y = "PDE Raw Score (nonwords read correctly)",
       caption = "excludes 3 subjects with NAs") +
  theme_classic()

p_swe_pde_2

```

\newpage
## W-J Letter-Word ID

The W-J raw score is the number of items answered correctly before reaching the ceiling rule (6 highest items that were administered are incorrect)

```{r w-j plots, echo=FALSE, message=FALSE}

# TO-DO: Scatter (x = Age, y = Raw Score)

p_wj_1 <- ggplot(b2_clean, aes(x = age, y = wj_raw)) + 
  geom_point(aes(color = grade)) +
  geom_smooth(method = "lm", se = T, linetype = "dashed", color = "grey") + 
  labs(title = "W-J Letter-Word ID Raw Scores", 
       x = "Age",
       y = "W-J Raw Score (# correct items)") +
  scale_x_continuous(breaks = seq(5, 9, by=0.5)) + # adjusting x axis
  theme_classic()

p_wj_1

# TO-DO: Density

p_wj_2 <- ggplot(b2_clean, aes(x = wj_raw, fill = grade)) +
  #geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.3) +
  #geom_histogram(aes(y = ..density..), color = "black", fill = "white") +
  geom_density(alpha = 0.3) +
  geom_vline(data = b2_clean_means, 
             aes(xintercept = wj_raw, color = grade),
             linetype = "dashed") +
  labs(title = "W-J Raw Scores",
       x = "Raw Score (# correct)",
       y = "Density",
       caption = "Dotted lines indicate raw score means by grade") +
  theme_classic()

p_wj_2


```

\newpage
# Comparing T1 and T2

Pasting initial T2 corrplot stuff here for now

TO-DO: T1-T2 correlations

```{r correlations, echo=FALSE}

p_t2_corr <- b2_clean %>% 
  select(-c(pid, grade, age_year, swe_age, swe_grade, pde_age, pde_grade)) %>% 
  mutate(ctopp_c_raw = ctopp_c_raw * -1,
         ctopp_l_raw = ctopp_l_raw * -1) %>% 
  cor(use = "complete.obs") %>% 
  corrplot(method = "color")

p_both_corr <- b_both %>% 
  select(-c(pid, grade, age_year, t1.age_year)) %>% 
  mutate(ctopp_c_raw = ctopp_c_raw * -1,
         ctopp_l_raw = ctopp_l_raw * -1,
         t1.ctopp_c_raw = t1.ctopp_c_raw * -1,
         t1.ctopp_l_raw = t1.ctopp_l_raw * -1,) %>% 
  cor(use = "complete.obs") %>% 
  corrplot(method = "color")

```

```{r paired plots, echo=FALSE}

# trying out paired plots 

p_wj_pair_all <- b_both %>%
  rename(T1_WJ = t1.wj_raw, T2_WJ = wj_raw) %>% 
  ggpaired(b_both, cond1 = "T1_WJ", cond2 = "T2_WJ",
           color = "condition",
           line.color = "gray", 
           line.size = 0.4,
           facet.by = "grade",
           xlab = "Time Point", 
           ylab = "Raw Score (# correct)",
           title = "W-J LWID Raw Scores by Grade",
           legend = "none")

p_wj_pair_all

p_swe_pair_all <- b_both %>%
  rename(T1_SWE = t1.swe_raw, T2_SWE = swe_raw) %>% 
  ggpaired(b_both, cond1 = "T1_SWE", cond2 = "T2_SWE",
           color = "condition",
           line.color = "gray", 
           line.size = 0.4,
           facet.by = "grade",
           xlab = "Time Point", 
           ylab = "Raw Score (# correct)",
           title = "TOWRE SWE Raw Scores by Grade",
           legend = "none")

p_swe_pair_all

p_pde_pair_all <- b_both %>%
  rename(T1_PDE = t1.pde_raw, T2_PDE = pde_raw) %>% 
  ggpaired(b_both, cond1 = "T1_PDE", cond2 = "T2_PDE",
           color = "condition",
           line.color = "gray", 
           line.size = 0.4,
           facet.by = "grade",
           xlab = "Time Point", 
           ylab = "Raw Score (# correct)",
           title = "TOWRE PDE Raw Scores by Grade",
           legend = "none")

p_pde_pair_all

```

